### 2021-11-26 input/proposal.tex
reworkd proof of detection

\begin{theorem}[\label{theo:efb}\underline{F}orwarding of \underline{B}est (FB)
    $\implies\eventually$CP]
    % 
    Assuming reliable communication links ($s_{xy}(m) \iff \eventually
    r_{yx}(m)$), and a total order on messages ($<$), \processes
    eventually reach consistent partitioning if each \process delivers
    its best message among \underline{r}eceived messages $R_x$, and
    forwards its best message among \underline{d}elivered messages
    $D_x$.
    % , \ie
    % the message $\alpha_s^d$ with the smallest distance $d$ to the
    % source $s$
    % ($(\alpha_s^d = \min R_x) \implies (d_x(\alpha_s^d) \wedge
    % f_x(\TODO{\alpha_s^d}))$).
\end{theorem}

@@

\begin{definition}[\label{def:echo}FB$^+$: echos] FB where a
  \process $x$ that receives a stale message sends back --~or
  echoes~-- its best up-to-date delivered message: $r_{xy}(\alpha_s)
  \wedge \delta_s \in D_x \implies s_{xy}(\alpha_{s'})$ with
  $\alpha_{s'} = \min D_x \wedge \delta_{s'} \not\in D_x$.
\end{definition}

\begin{lemma}[\label{def:fbfse}FB$^+ \wedge$ FS $\centernot\implies$ DCP]
  Forwarding best up-to-date, forwarding staleness, and echoing is not
  sufficient to guarantee dynamic consistent partitioning.
\end{lemma}

@@

\begin{lemma}[\label{def:detection}$\textup{FB}^+ \wedge \textup{FS}
  \implies$ \underline{D}etector existence (D)]
  Assuming $\textup{FB}^+$, $\textup{FS}$, a chain of delivery on
  $\pi_{xz}$ of $\alpha_x$ with $\alpha_x = \min D_z$, if $x$
  broadcasts a staleness notification $\delta_x$, either $z$
  eventually delivers it, or a \process $y\in\pi_{xz}$ eventually
  detects its blocking.
  % $b_p(\delta_x) \implies d_q(x) U Detect[X]$ ??? 
\end{lemma}

@@

  \noindent (B) Assuming a chain of delivery
  $\pi_{xz} = [x, \ldots, z]$ of $\alpha_x$ with at least
  $\alpha_x = \min D_z$, and assuming $d_x(\delta_x)$, then by (A),
  the staleness notification reaches a \process $\pi_{xz}[k]$. By
  Definition~\ref{def:fs}, the fowarding stops when $\pi_{xz}[k]$
  already delivered $\delta_x$ (covered by (A)) or delivered a better
  message $\alpha_x \neq \alpha_y = \min D_{\pi_{xz}[k]}$. The latter
  means that $\eventually d_{\pi_{xz}[k+1]}(\alpha_y)$ or
  $\delta_y \in D_{\pi_{xz}[k+1]}$.
  %
  %% (i) If every \process (such as $a$) starting from the source
  %% delivers and forwards a removal notification $\delta_X$ when their
  %% last delivery is $\alpha_X$, then every such \process eventually
  %% delivers the removal notification $\delta_X$ except \processes (such
  %% as $b$) that delivered $\alpha_X$ from a \process that delivered a
  %% message from another partition $\alpha_Y$ since then.
  %
  %% These exceptions eventually receive and deliver $\alpha_Y$ since
  %% $\alpha_Y$ is better than $\alpha_X$ through this path, except if
  %% they already received and delivered the removal notification of
  %% $P_Y$ (such as $c$). These \processes may never receive hence
  %% deliver $\delta_X$, and may never receive hence deliver a better
  %% message than $\alpha_X$. They need an additional mechanism to
  %% eventually purge $\alpha_X$ that cannot rely on the eventual purging
  %% of $P_Y$ at \processes like $b$, to avoid deadlocks.
  %
  %% (ii) Assuming that every \process keeps an history of its past
  %% deliveries, \processes (such as $c$) can detect the inconsistency,
  %% since they receive from their parent an already deleted partition
  %% $P_Y$. This notification means that the parent discards any
  %% $\delta_Z$ with $\alpha_Z^z < \alpha_Y^y$, and most importantly, if
  %% $\delta_X$ exists, it discards it. To ensure the purge of stale
  %% messages, a detecting \process must assume the worst case that such
  %% $\delta_X$ exists, and send another kind of message, noted $\Delta$,
  %% that notify the possible removal of $P_X$.
  %
  %% (iii) A \process (such as $d$) whose last delivery is $\alpha_X$,
  %% but whose parent is neither inconsistent (like $b$) nor receiving
  %% $\delta_X$ (like $a$), eventually receives $\Delta$ from a detecting
  %% \process, either directly or transitively, for such a child \process
  %% ($d$) trusts the possible removal of $P_X$ by delivering and
  %% forwarding such $\Delta$. $\Delta$ suffers from identical blocking
  %% conditions (between $b$ and $c$) than $\delta$, leading to the same
  %% solutions of detection and propagation of $\Delta$. Eventually,
  %% every \process whose last delivery is $\alpha_X$ receives and
  %% delivers either a better $\alpha_Z$, or $\delta_X$, or $\Delta_X$.
\end{proof}

>>>>

\begin{theorem}[\label{theo:fb}\underline{F}orwarding of \underline{B}est (FB)
    $\implies\eventually$CP]
    % 
    Assuming reliable communication links ($s_{xy}(m) \iff \eventually
    r_{yx}(m)$), and a total order on messages ($<$), \processes
    eventually reach consistent partitioning if each \process delivers
    its best message among \underline{r}eceived messages $R_x$, and
    forwards its best message among \underline{d}elivered messages
    $D_x$.
    % , \ie
    % the message $\alpha_s^d$ with the smallest distance $d$ to the
    % source $s$
    % ($(\alpha_s^d = \min R_x) \implies (d_x(\alpha_s^d) \wedge
    % f_x(\TODO{\alpha_s^d}))$).
\end{theorem}

@@

\begin{definition}[\label{def:fbplus}FB$^+$: echos] FB where a
  \process $x$ that receives a stale message sends back --~or
  echoes~-- its best up-to-date delivered message: $r_{xy}(\alpha_s)
  \wedge \delta_s \in D_x \implies s_{xy}(\alpha_{s'})$ with
  $\alpha_{s'} = \min D_x \wedge \delta_{s'} \not\in D_x$.
\end{definition}

\begin{lemma}[\label{lem:fbfse}FB$^+ \wedge$ FS $\centernot\implies$ DCP]
  Forwarding best up-to-date, forwarding staleness, and echoing is not
  sufficient to guarantee dynamic consistent partitioning.
\end{lemma}

@@

\begin{lemma}[\label{def:detection}$\textup{FB}^+ \wedge \textup{FS}
  \implies$ \underline{D}etector existence (D)]
  Assuming $\textup{FB}^+$, $\textup{FS}$, a chain of delivery on
  $\pi_{xz}$ of $\alpha_x$ with $\alpha_x = \min D_z$, if $x$
  broadcasts a staleness notification $\delta_x$, either $z$
  eventually delivers it, or a \process $y$ in such a chain of
  delivery $\pi_{xz}'$ eventually detects its blocking.
  % $b_p(\delta_x) \implies d_q(x) U Detect[X]$ ??? 
\end{lemma}

@@

  \noindent (B) Assuming a chain of delivery
  $\pi_{xz} = [x, \ldots, z]$ of $\alpha_x$ with at least
  $\alpha_x = \min D_z$, and assuming $d_x(\delta_x)$, then by (A),
  the staleness notification reaches a \process $ y = \pi_{xz}[k]$. By
  Definition~\ref{def:fs}, the fowarding stops when $y$ already
  delivered $\delta_x$ ($\delta_x \in D_y$ covered by (A)) or
  delivered a better message $\alpha_s = \min D_y$ with
  $\alpha_s \neq \alpha_x$. With $y'=\pi_{xz}[k+1]$, forwarding of
  best (Definition~\ref{def:fbplus}) implies three possible outcomes:
  \begin{enumerate}[(i)]
  \item $\alpha_{s'} = \min D_{y'}$ with $\alpha_{s'} \neq \alpha_x$
    hence $y' \iff y$ which leaves two possible outcomes:
  \item
    $r_{y'}(\alpha_s) \wedge \alpha_x = \min D_{y'} \wedge \alpha_x <
    \alpha_s$ which means that a shorter path of delivery $\pi_{xz}'$
    exists that either forwards appropriate staleness notifications
    (A) or detects its possible blocking:
  \item
    $r_{y'}(\alpha_s) \wedge \delta_s \in D_{y'} \wedge \alpha_x =
    \min D_{y'}$ which detects a possible blocking of
    $\delta_x$. Without global knowledge, $y'$ assumes it belongs to
    the shortest and only path of delivery of $\alpha_x$
    ($\pi_{xy'} \subseteq \Pi_{xz}$) thus it does not further delegate
    the detection to another \process.
  \end{enumerate} \vspace{-1.5em} %% ugly vspace but necessary for enumerate+qed
  %
  %% (i) If every \process (such as $a$) starting from the source
  %% delivers and forwards a removal notification $\delta_X$ when their
  %% last delivery is $\alpha_X$, then every such \process eventually
  %% delivers the removal notification $\delta_X$ except \processes (such
  %% as $b$) that delivered $\alpha_X$ from a \process that delivered a
  %% message from another partition $\alpha_Y$ since then.
  %
  %% These exceptions eventually receive and deliver $\alpha_Y$ since
  %% $\alpha_Y$ is better than $\alpha_X$ through this path, except if
  %% they already received and delivered the removal notification of
  %% $P_Y$ (such as $c$). These \processes may never receive hence
  %% deliver $\delta_X$, and may never receive hence deliver a better
  %% message than $\alpha_X$. They need an additional mechanism to
  %% eventually purge $\alpha_X$ that cannot rely on the eventual purging
  %% of $P_Y$ at \processes like $b$, to avoid deadlocks.
  %
  %% (ii) Assuming that every \process keeps an history of its past
  %% deliveries, \processes (such as $c$) can detect the inconsistency,
  %% since they receive from their parent an already deleted partition
  %% $P_Y$. This notification means that the parent discards any
  %% $\delta_Z$ with $\alpha_Z^z < \alpha_Y^y$, and most importantly, if
  %% $\delta_X$ exists, it discards it. To ensure the purge of stale
  %% messages, a detecting \process must assume the worst case that such
  %% $\delta_X$ exists, and send another kind of message, noted $\Delta$,
  %% that notify the possible removal of $P_X$.
  %
  %% (iii) A \process (such as $d$) whose last delivery is $\alpha_X$,
  %% but whose parent is neither inconsistent (like $b$) nor receiving
  %% $\delta_X$ (like $a$), eventually receives $\Delta$ from a detecting
  %% \process, either directly or transitively, for such a child \process
  %% ($d$) trusts the possible removal of $P_X$ by delivering and
  %% forwarding such $\Delta$. $\Delta$ suffers from identical blocking
  %% conditions (between $b$ and $c$) than $\delta$, leading to the same
  %% solutions of detection and propagation of $\Delta$. Eventually,
  %% every \process whose last delivery is $\alpha_X$ receives and
  %% delivers either a better $\alpha_Z$, or $\delta_X$, or $\Delta_X$.
\end{proof}
